An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-04-12 EconGrowth-R-Functions.Rnw
Now doing CES without energy in the cesData function.

###############################
# Tells whether we have data for this combination of country and energy type for a Cobb-Douglas curve fit.
# a missing energyType or energyType == NA means that we want to develop a model without energy, if possible.
#
# returns logical value (TRUE or FALSE)
##
haveDataCD <- function(countryAbbrev, energyType){
  # The design here is to go through several cases where data is unavailable and
  # return FALSE. If we make it through all of those cases, we'll return TRUE.
  if (!(countryAbbrev %in% countryAbbrevs)){
    # If we don't know the country, we don't have data
    return(FALSE)
  }
  if (missing(energyType) || is.na(energyType)){
    # If we don't want energy included, we have data for all known countries. We need to check this case
    # first, BEFORE doing various checks on the energyType.
    return(TRUE)
  }
  if (!(energyType %in% energyTypes)){
    # If we want to include energy, we know only a few types. So, return false if we don't know the energy type.
    return(FALSE)
  }
  if (energyType == "U" && (!(countryAbbrev %in% countryAbbrevsU))){
    # We have U data for only a few countries.
    return(FALSE)
  }
  return(TRUE)
}

@@

######################################
# Loads data and trims rows where necessary.
# Also, replaces the name of the desired energy column with "iEToFit".
# If you omit the energyType parameter, "iEToFit" will not appear in the column names, because
# it is assumed that you want a CES fit without an energy term.
# Returns a data.frame with the loaded data.
##
loadAndPrepDataForCES <- function(countryAbbrev, energyType){
  dataTable <- loadData(countryAbbrev)
  # Trim data from China and South Africa that is missing labor information for 2011.
  dataTable <- subset(dataTable, !is.na(iLabor))
  if (!missing(energyType)){
    if (energyType == "U"){
      # Trim the dataset to include only those years for which U is available.
      dataTable <- subset(dataTable, !is.na(iU))
    }
    # We need to do the CES fit with the desired energyType.
    # To achieve the correct fit, we'll change the name of the desired column
    # to "iEToFit" and use "iEToFit" in the nls function.
    dataTable <- replaceColName(dataTable, energyType, "iEToFit")
  }
  return(dataTable)
}

@@

@@

@@

# returns a data.frame of data for the LINEX model.

@@

cesData <- function(countryAbbrev, energyType){

@@

@@

#print(summaryCES)

@@

@@

#print(dofCES)

@@

#print(tvalCES)

@@

#print(ciCES)
  # Get confidence intervals for the elasticities of substitution
  sigma_1.est <- deltaMethod(modelCES, "1.0 / (1.0 + rho_1)") # Estimates sigma_1 and its standard error (SE).
#print(sigma_1.est)

@@

  # There are some cases where the confidence intervals for sigma_1 are huge. Set them to -Inf or Inf as appropriate
  if (is.nan(sigma_1CI[1]) || (sigma_1CI[1] < -1.0e10)){
    sigma_1CI[1] <- -Inf 

@@

  if (is.nan(sigma_1CI[2]) || (sigma_1CI[2] > 1.0e10)){
    sigma_1CI[2] <- Inf
#print("sigma_1CI:")
#print(sigma_1CI)
#print(sigma.est)
#print("sigmaCI:")
#print(sigmaCI)
  gamma <- coef(modelCES)["gamma"] #as.numeric(coef(modelCES)["gamma"])
  lambda <- coef(modelCES)["lambda"] #as.numeric(coef(modelCES)["lambda"])

@@

#print(estCES)

@@

#print(ciCES["gamma","2.5 %"])

@@

#print(lowerCES)

@@

#print(upperCES)

@@

#print("dataCES")
#print(dataCES)

>>>>

###############################
# Tells whether we have data for this combination of country and energy type for a Cobb-Douglas curve fit.
# a missing energyType or energyType == NA means that we want to develop a model without energy, if possible.
#
# returns logical value (TRUE or FALSE)
##
haveDataCD <- function(countryAbbrev, energyType){
  # The design here is to go through several cases where data is unavailable and
  # return FALSE. If we make it through all of those cases, we'll return TRUE.
  if (!(countryAbbrev %in% countryAbbrevs)){
    # If we don't know the country, we don't have data
    return(FALSE)
  }
  if (missing(energyType) || is.na(energyType) || is.null(energyType)){
    # If we don't want energy included, we have data for all known countries. We need to check this case
    # first, BEFORE doing various checks on the energyType.
    return(TRUE)
  }
  if (!(energyType %in% energyTypes)){
    # If we want to include energy, we know only a few types. So, return false if we don't know the energy type.
    return(FALSE)
  }
  if (energyType == "U" && (!(countryAbbrev %in% countryAbbrevsU))){
    # We have U data for only a few countries.
    return(FALSE)
  }
  return(TRUE)
}

###############################
# Tells whether we have data for this combination of country and energy type for a CES curve fit.
# a missing energyType or energyType == NA means that we want to develop a model without energy, if possible.
# This function re-routes to haveDataCD, because the logic is identical.
#
# returns logical value (TRUE or FALSE)
##
haveDataCES <- function(countryAbbrev, energyType){
  return(haveDataCD(countryAbbrev, energyType))
}

###############################
# Tells whether we have data for this combination of country and energy type for a LINEX curve fit.
# a missing energyType or energyType == NA means that we want to develop a model without energy, if possible.
# This function re-routes to haveDataCD, because the logic is identical.
#
# returns logical value (TRUE or FALSE)
##
haveDataLINEX <- function(countryAbbrev, energyType){
  return(haveDataCD(countryAbbrev, energyType))
}

@@

######################################
# Loads data and trims rows where necessary.
# Also, replaces the name of the desired energy column with "iEToFit".
# If you omit the energyType parameter, "iEToFit" will not appear in the column names, because
# it is assumed that you want a CES fit without an energy term.
# Returns a data.frame with the loaded data.
##
loadAndPrepDataForCES <- function(countryAbbrev, energyType = NULL){
  dataTable <- loadData(countryAbbrev)
  # Trim data from China and South Africa that is missing labor information for 2011.
  dataTable <- subset(dataTable, !is.na(iLabor))
  if (missing(energyType) || is.na(energyType) || is.null(energyType) || energyType == ""){
    # We're done. Don't have to deal with special issues with energyType.
    return(dataTable)
  }
  # We have an energyType argument. Do some additional checking.
  if (energyType == "U"){
    # Trim the dataset to include only those years for which U is available.
    dataTable <- subset(dataTable, !is.na(iU))
  }
  # We need to do the CES fit with the desired energyType.
  # To achieve the correct fit, we'll change the name of the desired column
  # to "iEToFit" and use "iEToFit" in the nls function.
  dataTable <- replaceColName(dataTable, energyType, "iEToFit")
  return(dataTable)
}

@@

  if (missing(energyType) || is.null(energyType) || is.na(energyType) || energyType == ""){

@@

# energyType is a string, one of "Q", "X", or "U". Not specifying an energy type means we're interested
# in a CES function without energy.

@@

# returns a data.frame of data for the CES model.

@@

cesData <- function(countryAbbrev, energyType=NULL){

@@

  if (!haveDataCES(countryAbbrev, energyType)){
    #Return a column of NAs if we don't have data available for this combination of country and energy type.

@@

@@

  if (is.null(energyType)){
    # For a CES model without energy, we have 4 parameters: gamma, lambda, delta, and rho
    dofCES <- nrow(loadAndPrepDataForCES(countryAbbrev)) - 4 #    
    # For a CES model with energy, we have 6 parameters: gamma, lambda, delta, rho, delta_1, and rho_1

@@

@@

@@

  # Get confidence intervals for the elasticities of substitution.
  # Note: I would rather get the value of sigma and its standard error directly from the cesModel object.
  # The information is all in there. You can see it as the E terms at the bottom of the
  # print(summaryCES) results.
  if (!is.null(energyType)){
    sigma_1.est <- deltaMethod(modelCES, "1.0 / (1.0 + rho_1)") # Estimates sigma_1 and its standard error (SE)

@@

@@

  if (is.null(energyType)){
    # We're not using energy in this fit.
    gamma <- coef(modelCES)["gamma"]
    lambda <- coef(modelCES)["lambda"]
    # Without energy, we get the delta coefficient that means the split between 
    # capital and labor.
    # With energy, delta_1 means the split between energy and labor.
    # To be consistent, we map the delta from the "without energy" model to variable name "delta_1"
    delta_1 <- coef(modelCES)["delta"]
    # The without energy model can be obtained with delta = 1.0. So, we set it here.
    delta <- 1.0
    # The without energy model has a sigma (elasticity of substitution between k and l) that is 
    # equivalent to the sigma_1 parameter from the "with energy" model. So, we map the names 
    # here.
    sigma_1 <- sigma.est$Estimate
    # With delta = 1.0, rho has no meaning (there is no energy term). Therefore, sigma has no meaning.
    # So, set sigma to NA
    sigma <- NA
    # The size of the CI on delta is 0.0 for the CES model without energy, so we set the upper and lower bounds to 1.0
    # The CI on delta_1 shows up with the variable delta for the CES model without energy.
                           delta_1=ciCES["delta", "2.5 %"], delta=1.0,
                           sigma_1=sigmaCI[1], sigma=NA)
                           delta_1=ciCES["delta", "97.5 %"], delta=1.0,
                           sigma_1=sigmaCI[2], sigma=NA)
    # We're using energy in this fit.
    gamma <- coef(modelCES)["gamma"]
    lambda <- coef(modelCES)["lambda"]

@@

@@

@@

@@

@@

### 2013-04-12 EconGrowth.Rnw
Now doing CES without energy in the cesData function.

@@

# print("US")
# print(summary(cesModel("US")))
# print("UK")
# print(summary(cesModel("UK")))
# print("JP")
# print(summary(cesModel("JP")))
# print("CN")
# print(summary(cesModel("CN")))
# print("ZA")
# print(summary(cesModel("ZA")))
# print("SA")
# print(summary(cesModel("SA")))
# print("IR")
# print(summary(cesModel("IR")))
# print("TZ")
# print(summary(cesModel("TZ")))
# print("ZM")
# print(summary(cesModel("ZM")))

# print(cesPredictionsColumn("Q"))
# print(cesPredictionsColumn())

# createCESLatticeGraph()

>>>>

      \item{Get values and se for $\sigma$s directly from the \texttt{cesModel} object in the \texttt{cesData} function}

@@

print(cesData("US"))
